Spatiotemporal Multi-Resolution Approximation of the Amari Type Neural Field Model

### 2012-02-12 Multiresolution_framework/ltx/manuscript.tex
minor edits

To demonstrate the performance of the MRAIDE estimation framework, data was generated synthetically  using Eqs.~\eqref{eq:DiscreteTimeModel} and \eqref{eq:ObservationEquation}, allowing a comparison between true and estimated parameters. In doing so, the membrane time constant was set to $\tau = 10$~ms \citep{David2003}, and the sampling time was chosen ten times faster \citep{Stephan2008}, i.e. $T_s = 1$~ms, ensuring the discrepancy between the continuous neural field model and its discrete approximation is small. The connectivity kernel used for data generation composed of two B-spline scaling functions at $j=0$ and $j=1$ forming a compact support Mexican hat shape (see Fig.\ref{fig:KernelEstimate}) given by

@@

with $\theta_1=200$ and $\theta_2=-100$. \dean{This is depicted in Figure~\ref{fig:}}

@@

\section{Lower bound on likelihood function}\label{ap:Lowerbound}
A derivation of the lower bound on the likelihood function that is based on \citet{Bishop2006} is provided in this Appendix. We want to maximize the likelihood function, 
\begin{equation}
 p(\mathbf Y;\boldsymbol\theta)=\int_{\mathbf X}p(\mathbf X,\mathbf Y;\boldsymbol\theta)\mathrm{d}\mathbf X,
\end{equation}
which is difficult to perform directly. Maximizing the likelihood in terms of the parameter set is equivalent to the log-likelihood maximization, as both the likelihood and log-likelihood functions share the same maximizer due to the monotonicity of the logarithm function. For any choice of distribution defined over the latent variables, $q(\mathbf X)$, the following decomposition holds
\begin{equation}\label{eq:EM-loglikelihooddecomposition}
 \ln~p(\mathbf Y;\boldsymbol\theta)=\mathcal{L}(q,\boldsymbol\theta)+\mathrm{KL}(q\parallel p),
\end{equation}
where
\begin{align}
 \mathcal{L}(q,\boldsymbol\theta)&=\int_{\mathbf X} q(\mathbf X)\ln\left\lbrace \frac{p(\mathbf X,\mathbf Y;\boldsymbol\theta)}{q(\mathbf X)} \right\rbrace \mathrm{d}\mathbf X \label{eq:EM-decomposition1}\\
\mathrm{KL}(q\parallel p)&=-\int_{\mathbf X}q(\mathbf X)\ln\left\lbrace \frac{p(\mathbf X \mid \mathbf Y;\boldsymbol\theta)}{q(\mathbf X)} \right\rbrace \mathrm{d}\mathbf X. \label{eq:EM-decomposition2}
\end{align}
Eq.~\eqref{eq:EM-loglikelihooddecomposition} can be verified by applying the probability product rule to the joint distribution $p(\mathbf X,\mathbf Y;\boldsymbol\theta)$ in Eq.~\eqref{eq:EM-decomposition1} giving
\begin{align}
	\mathcal{L}(q,\boldsymbol\theta)+\mathrm{KL}(q\parallel p)&= \int_{\mathbf X} q(\mathbf X)\ln\left\lbrace \frac{p(\mathbf X \mid \mathbf Y;\boldsymbol\theta)p(\mathbf Y;\boldsymbol\theta)}{q(\mathbf X)} \right\rbrace \mathrm{d}\mathbf X \nonumber\\
	&- \int_{\mathbf X}q(\mathbf X)\ln\left\lbrace \frac{p(\mathbf X \mid \mathbf Y;\boldsymbol\theta)}{q(\mathbf X)} \right\rbrace \mathrm{d}\mathbf X\nonumber\\
	&= \int_{\mathbf X}q(\mathbf X)\ln p(\mathbf Y; \boldsymbol\theta)\mathrm{d}\mathbf X.
\end{align}
Noting  $q(\mathbf X)$ is a normalized distribution that integrates to one gives the log-likelihood function. We also note that Eq.~\eqref{eq:EM-decomposition2} is the Kullback-Liebler distance between distributions $p(\mathbf X \mid \mathbf Y;\boldsymbol\theta)$ and $q(\mathbf X)$. Therefore,
\begin{equation}
	\mathrm{KL}(q\parallel p)\ge 0.
\end{equation}   
The equality holds if and only if $q(\mathbf X)=p(\mathbf X \mid \mathbf Y;\boldsymbol\theta)$. It follows from  Eq.~\eqref{eq:EM-loglikelihooddecomposition}, that the quantity $\mathcal{L}(q,\boldsymbol\theta)$ is a lower bound on the log-likelihood function, i.e.
\begin{equation}
 \mathcal{L}(q,\boldsymbol\theta)\le \ln~p(\mathbf Y\mid\boldsymbol\theta)
	\end{equation}

@@

The expressions for $\boldsymbol\Xi_{0}$ and $\boldsymbol\Xi_{1}$ in terms of smoother outputs can be found in \cite{Shumway2000}, here we present the derivation for completeness.  The expression for  $\boldsymbol\Xi_{0}$ as defined in Eq.~\eqref{eq:MRA-Xi0} of the main text is given by

>>>>

To demonstrate the performance of the MRAIDE estimation framework, data was generated synthetically  using Eqs.~\eqref{eq:DiscreteTimeModel} and \eqref{eq:ObservationEquation}, allowing a comparison between true and estimated parameters. In doing so, the membrane time constant was set to $\tau = 10$~ms \citep{David2003}, and the sampling time was chosen ten times faster \citep{Stephan2008}, i.e. $T_s = 1$~ms, ensuring the discrepancy between the continuous neural field model and its discrete approximation is small. The connectivity kernel used for data generation composed of two B-spline scaling functions at $j=0$ and $j=1$ forming a compact support Mexican hat shape 

@@

with $\theta_1=200$ and $\theta_2=-100$. This is depicted in Fig.~\ref{fig:KernelEstimate}(A).

@@

\section{Lower bound on likelihood function}\label{ap:Lowerbound}
A derivation of the lower bound on the likelihood function that is based on \citet{Bishop2006} is provided in this Appendix. We want to maximize the likelihood function, 
\begin{equation}
 p(\mathbf Y;\boldsymbol\theta)=\int_{\mathbf X}p(\mathbf X,\mathbf Y;\boldsymbol\theta)\mathrm{d}\mathbf X,
\end{equation}
which is difficult to perform directly. Maximizing the likelihood in terms of the parameter set is equivalent to the log-likelihood maximization, as both the likelihood and log-likelihood functions share the same maximizer due to the monotonicity of the logarithm function. For any choice of distribution defined over the hidden variables, $q(\mathbf X)$, the following decomposition holds
\begin{equation}\label{eq:EM-loglikelihooddecomposition}
 \ln~p(\mathbf Y;\boldsymbol\theta)=\mathcal{L}(q,\boldsymbol\theta)+\mathrm{KL}(q\parallel p),
\end{equation}
where
\begin{align}
 \mathcal{L}(q,\boldsymbol\theta)&=\int_{\mathbf X} q(\mathbf X)\ln\left\lbrace \frac{p(\mathbf X,\mathbf Y;\boldsymbol\theta)}{q(\mathbf X)} \right\rbrace \mathrm{d}\mathbf X \label{eq:EM-decomposition1}\\
\mathrm{KL}(q\parallel p)&=-\int_{\mathbf X}q(\mathbf X)\ln\left\lbrace \frac{p(\mathbf X \mid \mathbf Y;\boldsymbol\theta)}{q(\mathbf X)} \right\rbrace \mathrm{d}\mathbf X. \label{eq:EM-decomposition2}
\end{align}
Eq.~\eqref{eq:EM-loglikelihooddecomposition} can be verified by applying the probability product rule to the joint distribution $p(\mathbf X,\mathbf Y;\boldsymbol\theta)$ in Eq.~\eqref{eq:EM-decomposition1} giving
\begin{align}
	\mathcal{L}(q,\boldsymbol\theta)+\mathrm{KL}(q\parallel p)&= \int_{\mathbf X} q(\mathbf X)\ln\left\lbrace \frac{p(\mathbf X \mid \mathbf Y;\boldsymbol\theta)p(\mathbf Y;\boldsymbol\theta)}{q(\mathbf X)} \right\rbrace \mathrm{d}\mathbf X \nonumber\\
	&- \int_{\mathbf X}q(\mathbf X)\ln\left\lbrace \frac{p(\mathbf X \mid \mathbf Y;\boldsymbol\theta)}{q(\mathbf X)} \right\rbrace \mathrm{d}\mathbf X\nonumber\\
	&= \int_{\mathbf X}q(\mathbf X)\ln p(\mathbf Y; \boldsymbol\theta)\mathrm{d}\mathbf X.
\end{align}
Noting  $q(\mathbf X)$ is a normalized distribution that integrates to one gives the log-likelihood function. We also note that Eq.~\eqref{eq:EM-decomposition2} is the Kullback-Liebler distance between distributions $p(\mathbf X \mid \mathbf Y;\boldsymbol\theta)$ and $q(\mathbf X)$. Therefore,
\begin{equation}
	\mathrm{KL}(q\parallel p)\ge 0.
\end{equation}   
The equality holds if and only if $q(\mathbf X)=p(\mathbf X \mid \mathbf Y;\boldsymbol\theta)$. It follows from  Eq.~\eqref{eq:EM-loglikelihooddecomposition}, that the quantity $\mathcal{L}(q,\boldsymbol\theta)$ is a lower bound on the log-likelihood function, i.e.
\begin{equation}
 \mathcal{L}(q,\boldsymbol\theta)\le \ln~p(\mathbf Y\mid\boldsymbol\theta)
	\end{equation}

@@

The expressions for $\boldsymbol\Xi_{0}$ and $\boldsymbol\Xi_{1}$ in terms of smoother outputs can be found in \cite{Shumway2000}, here we present the derivation for completeness.  The expression for  $\boldsymbol\Xi_{0}$ as defined in Eq.~\eqref{eq:app-Xi0} of the main text is given by
